Visualization and analysis of probability distributions of large temporal data

### 2022-02-17 report/response.Rmd

> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the distribution of each vt.

To do.

> Q10. Where is $\mu$ in Table 4.2?

Done.
Clarification: $\mu$ is the difference between means considered for consecutive categories.

@@

> Q13. Figure 4.7 is missing labels (a), (b) (c).

Done. Added a, b representing 12 customers (rows) in each of them.

> Q14. Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?

Done.
The caption and description of the figure in the text have been changed to indicate that Fig. 4.9 depicts a visual summary of why three clusters are insufficient for this data set, however it is recommended by sindex summary plots. In a cluster analysis, it is standard practice to test a few cluster sizes before and after the optimal $k$ to see if a different number of clusters is useful. In this example, increasing the number of clusters actually unfolds some behavior which was not apparent with $k=3$ as suggested by sindex.

### 2022-02-17 Rmd/04-gracsr.Rmd
Ch4 Q14 Catherine

<!--Figure \ref{fig:opt-cluster-tsne-jsd} The *sindex* is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}. The *sindex* falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases. In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters from JS-NQT clustering in (a) and (b) respectively, and helps to characterize each cluster. In the quantile plots the line represents the median, and the region shows the area between the $25^{th}$ and $75^{th}$ percentiles. The only difference between the four and five cluster solution is that A-4 divides further into B-4 and B-5. This additional division makes a clearer clustering, because it resolves the heterogeneity in *moy* creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage. B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive *hod* patterns but are both heterogeneous in *moy* and *wnwd*. B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on *hod* and *wnwd*. This clustering almost agrees with the clusters visible in the t-SNE plot.

@@

(ref:groups-4and5) Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over *hod*, *moy*, and *wnwd*. Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively. Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the *moy* distribution.

>>>>

<!--Figure \ref{fig:opt-cluster-tsne-jsd} The *sindex* is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}. The *sindex* falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases. In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters from JS-NQT clustering in (a) and (b) respectively, and helps to characterize each cluster. In the quantile plots the line represents the median, and the region shows the area between the $25^{th}$ and $75^{th}$ percentiles. The only difference between the four and five cluster solution is that A-4 divides further into B-4 and B-5. This additional division makes a clearer clustering, because it resolves the heterogeneity in *moy* creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage. B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive *hod* patterns but are both heterogeneous in *moy* and *wnwd*. B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on *hod* and *wnwd*. This clustering almost agrees with the clusters visible in the t-SNE plot. This display also serves as a visual summary of why three clusters are insufficient using JS-NQT approach, however it is recommended by *sindex* summary plots in Figure \ref{fig:opt-cluster-tsne-jsd}.

@@

(ref:groups-4and5) Visual summary of why three clusters are insufficient for JS-NQT approach, however it is recommended by *sindex*. The plot shows four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over *hod*, *moy*, and *wnwd*. Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively. Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the *moy* distribution.
